A Large-Scale Study of the Use of Eval in R

### 2020-08-27 inst/paper/main.tex
More env examples

One of the main differences of the R \eval with the javascript one is how it
interacts with \emph{scope}, or environments in R. In Javascript, \c{eval}
can access the local scope and the global scope. In R, \c{eval} has an
argument \c{envir} to indicate in which environment its first argument,
after being resolved, must be evaluated. By default, it is equal to \c{parent.frame()}, which is the parent environment of the \c{eval} call. This default argument happens in \miss{XX\%} of the eval calls.

@@

There are two main environments affected by \eval:
\begin{itemize}
	\item the \emph{existing} environment where \c{eval} reads. It is the first non new environment. \c{eval} will also be able to read in all parent environment of that environment. It can be the global environment, a package environment, an empty environment, the callee environment, or a parent environment of the callee.
	\item the \emph{default} assignment environment \c{eval} writes new bindings.\footnote{It is the default one but the expression in \eval can always write in any other environments using \c{assign} or in the parent environment using \c{<<-}.} If it is different from the read environment and is a new one that inherits from the read environment, we add a symbol \texttt{+} after the read environment.
\end{itemize}

As shown in Figure~\ref{fig:environments}, \eval in R is often performed in the global environment.  It is also used a lot to evaluate in the context of a dataframe (\emph{list} environment).

% Variability: how many different envir class, and different envir expr, for one call site?
\begin{figure*}[htbp]
	\centering
	\includegraphics[width=\linewidth]{environments.pdf}
	\caption{The environments in which the expression is evaluated. \c{NA} corresponds to the default environment.} \label{fig:environments}
\end{figure*}

>>>>

One of the main differences of the R \eval with the javascript one is how it
interacts with \emph{scope}, or environments in R. In Javascript, \c{eval}
can access the local scope and the global scope. In R, \c{eval} has an
argument \c{envir} to indicate in which environment its first argument,
after being resolved, must be evaluated. By default, it is equal to \c{parent.frame()}, which is the parent environment of the \c{eval} call. This default argument happens in \DefaultEnvirExprPercent of the eval calls.

@@

There are two main environments affected by \eval:
\begin{itemize}
	\item the \emph{existing} environment where \c{eval} reads. It is the first non new environment. \c{eval} will also be able to read in all parent environment of that environment. It can be the global environment, a package environment, an empty environment, the callee environment, or a parent environment of the callee: $1$ is the parent, $2$ the grand-parent, and so on.
	\item the \emph{default} assignment environment \c{eval} writes new bindings.\footnote{It is the default one but the expression in \eval can always write in any other environments using \c{assign} or in the parent environment using \c{<<-}.} If it is different from the read environment and is a new one that inherits from the read environment, we add a symbol \texttt{+} after the read environment.
\end{itemize}

To give some examples,  global+ refers to a new scope that inherits from the global scope, while 3 refers to the great grand parent of the \eval scope.

Table~\ref{tab:environments} shows the environments when they are \emph{explicitly specified}. \eval in R is often performed in a new environment inherited from the global environment for packages, while for core, it is mostly done in the environment of the grand-parent.  It is also used a lot to evaluate in the context of a dataframe (\emph{list} environment). \emph{loop} \textbf{To explain...? But we do not see it here anyway}

The \c{sys.call} option is not used a lot, only \SysCallEnvirPercent as a whole, and always with \c{sys.parent()} as argument.

% Variability: how many different envir class, and different envir expr, for one call site?
%\begin{figure*}[htbp]
%	\centering
%	\includegraphics[width=\linewidth]{environments.pdf}
%	\caption{The environments in which the expression is evaluated. \c{NA} corresponds to the default environment.} \label{fig:environments}
%\end{figure*}

\begin{table}[htbp]
    \centering
    \begin{tabular}{|c|c||c|c|}
        \hline
        \multicolumn{2}{|c|}{Core} &  \multicolumn{2}{|c|}{Packages}  \\
        \hline
        Environment & Proportion & Environment & Proportion \\
        \hline
        \CoreEnvClassA &  \CoreProportionA & \PackagesEnvClassA &  \PackagesProportionA\\
        \hline
        \CoreEnvClassB &  \CoreProportionB & \PackagesEnvClassB &  \PackagesProportionB \\
        \hline
        \CoreEnvClassC & \CoreProportionC  & \PackagesEnvClassC &  \PackagesProportionC \\
        \hline
       \CoreEnvClassD  &  \CoreProportionD & \PackagesEnvClassD &  \PackagesProportionD \\
        \hline
        \CoreEnvClassE  & \CoreProportionE & \PackagesEnvClassE & \PackagesProportionE \\
        \hline
    \end{tabular} \label{tab:environments}
\caption{Environments in terms of eval calls.}
\end{table}

\c{match.arg} (see Listing~\ref{code:matchargenvir}) is the function that contributes the most to the eval calls (\MatchArgPercent) and has environment class 2. \c{sys.parent()} refers to the parent environment of \c{match.arg} and so to the grand-parent environment of \c{eval}, \ie what we denote 2.

\begin{lstlisting}[label=code:matchargenvir, caption={\c{match.arg} using \eval with a custom environment.}]
formal.args <- formals(sys.function(sysP <- sys.parent()))
choices <- eval(formal.args[[as.character(substitute(arg))]], envir = sys.frame(sysP))
\end{lstlisting}

In package \emph{gamlss}, function \c{get.K} evaluates its expression in its parent environment, the one of the body of \c{get.K} returned from \c{environment()}.

\begin{lstlisting}
eval(parse(text = paste(\"fam$dld\", ii, sep = \"\")), envir = environment())
\end{lstlisting}

The \emph{glue} library is used to interpolate string litterals with variables, such as in \c{glue('My name is {name}.')} where \c{name} will be subtituted by its value. \emph{glue} makes it possible to define transformers that process those interpolated variables. They are built upon the \c{identity_transformer}, which uses \eval. The interpretation environment for \eval is often global here, as the string litterals are often used to print variabels at the top level.

\begin{lstlisting}
identity_transformer <- function(text, envir) {
    eval(parse(text = text, keep.source = FALSE), envir)
}
\end{lstlisting}

The \c{imchange} function of package \c{imager} makes it possible to modify images using a dedicated formula syntax using \c{~}.\footnote{Inspired by \c{map} in package \emph{purr}.}  Here, \eval is evaluated in \c{newenv}, which creates a new environment that inherits from \c{parent.frame()} by default. We classify it as 1+.

\begin{lstlisting}
newenv <- new.env()
...
fo <- parse(text=as.character(fo)[2])
im[where] <- eval(fo,envir=newenv,enclos=env)
\end{lstlisting}
